---
title: "Connect to BigQuery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Connect to BigQuery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This guide walks you through getting access to the OpenPrescribing BigQuery database and connecting to it from R with `connect_bq()`.

## Get access

To get access to the OpenPrescribing BigQuery database write a message in the `#openprescribing` Slack channel and tag `bennett-admins`.
One of the admins will send you a JSON file with your BigQuery credentials in a private message.
The file should look something like this:

```json
{
  "type": "ACCOUNT_TYPE",
  "project_id": "PROJECT_ID",
  "private_key_id": "PRIVATE_KEY_ID",
  "private_key": "PRIVATE_KEY",
  "client_email": "CLIENT_EMAIL",
  "client_id": "CLIENT_ID",
  "auth_uri": "AUTH_URI",
  "token_uri": "TOKEN_URI",
  "auth_provider_x509_cert_url": "AUTH_CERTIFICATE",
  "client_x509_cert_url": "CLIENT_CERTIFICATE",
  "universe_domain": "DOMAIN"
}
```

**Never commit this file to GitHub or share it with anyone.**

## Store your credentials

You can store this file anywhere on your computer, but a common location is in a hidden folder in your home directory, for example:

**On macOS/Linux:**

```bash
~/.config/openprescribing/bq-credentials.json
```

**On Windows:**

```bash
C:/Users/YOUR_USERNAME/.config/openprescribing/bq-credentials.json
```

## Set up the environment variable

Rather than hardcoding the path in your R scripts you can set an environment variable named `OP_CREDENTIALS`.
To help you edit your `.Renviron` file and set the variable for all R sessions you can use the `usethis` package:

```{r, eval = FALSE}
# Open your .Renviron file
usethis::edit_r_environ()
```

This will open your `.Renviron` file in your editor where you can add the location where you stored your credentials file, for example:

```bash
# macOS/Linux
OP_CREDENTIALS=~/.config/openprescribing/bq-credentials.json

# Windows
OP_CREDENTIALS=C:/Users/YOUR_USERNAME/.config/openprescribing/bq-credentials.json
```

**Note: You will probably need to restart your R session for R to know about your new environment variable.**

## Connect from R

With the environment variable in place, `connect_bq()` authenticates and connects to the `hscic` dataset in the `ebmdatalab` project.
You might be asked to log in with your Google account the first time you run this.

```{r, eval = FALSE}
library(opr)

con <- connect_bq()
```

You can connect to a different dataset, or use a differently named environment variable, via the arguments:

```{r, eval = FALSE}
con <- connect_bq(dataset = "dmd", credentials_var = "MY_OTHER_CREDENTIALS")
```

The connection works with everything in the `DBI` and `dplyr`/`dbplyr` ecosystem, for example:

```{r, eval = FALSE}
# List all available tables
DBI::dbListTables(con)

# Create a lazy query against any table
dplyr::tbl(con, "practices")
```

If something is not set up correctly, `connect_bq()` will tell you what to fix, for example when the environment variable is missing:

```{r, error = TRUE}
opr::connect_bq(credentials_var = "VARIABLE_THAT_IS_NOT_SET")
```

## Next steps

See `vignette("opr")` for how to query the data with the `get_*()` functions, and `vignette("database-schema")` for descriptions of the main tables.
