---
title: "Database schema"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Database schema}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(opr)
```

This vignette describes the main tables used for research projects.
The `hscic` dataset has many more tables (around 194); to list them all, run:

```{r, eval = FALSE}
con <- connect_bq()
DBI::dbListTables(con)
```

Each table below comes with a five-row sample bundled in the package (for example `practices_sample`), so you can look at the data without a BigQuery connection.

## `normalised_prescribing`

A cleaned-up version of the main prescribing dataset, updated monthly.
This is the table you will usually want, rather than the raw prescribing table.
Query it with `get_normalised_prescribing()`.

| Column | Description |
|---|---|
| `sha` | Strategic Health Authority code (e.g. QHM) |
| `regional_team` | Regional team code (e.g. Y63) |
| `stp` | Sustainability and Transformation Partnership code (e.g. QHM) |
| `pct` | Primary Care Trust code (e.g. 16C) |
| `practice` | Practice code (e.g. A81001) |
| `bnf_code` | British National Formulary (BNF) code (e.g. 0408010W0BBAKA1) |
| `bnf_name` | British National Formulary (BNF) drug name (e.g. Epilim Chrono 300 tablets) |
| `items` | Number of items (one item = one appearance on a prescription) |
| `net_cost` | Cost according to the Drug Tariff |
| `actual_cost` | Actual cost after adjustments for bulk purchases, out of pocket expenses, etc. This is the cost field usually used in analyses |
| `quantity` | Number of pills/grams/millilitres/dressings/ampoules prescribed |
| `month` | Prescribing month |

```{r}
normalised_prescribing_sample
```

## `practices`

GP practice details including addresses and organisational information, updated monthly.
Contains the latest month's data only, but includes closed practices.
Query it with `get_practices()`.

| Column | Description |
|---|---|
| `code` | Practice code (e.g. E82084) |
| `name` | Practice name (e.g. HATFIELD ROAD SURGERY) |
| `address1` to `address5` | Address lines |
| `postcode` | Practice postcode (e.g. AL1 3HD) |
| `location` | Location of the practice in WGS84 format |
| `ccg_id` | Clinical Commissioning Group identifier (e.g. 06N) |
| `pcn_id` | Primary Care Network identifier (e.g. U73966) |
| `setting` | Practice setting code, e.g. 4 = GP Practice. See `practice_settings` for all codes |
| `close_date` | Date the practice closed |
| `join_provider_date` | Date the practice joined its CCG |
| `leave_provider_date` | Date the practice left its CCG (or closed) |
| `open_date` | Date the practice opened |
| `status_code` | Practice status code, e.g. A = Active. See `practice_statuses` for all codes |

```{r}
practices_sample
```

## `practice_statistics`

Practice-level patient counts by age and gender, updated monthly.
Contains the latest month's data only.
Also contains population weightings used to compare prescribing across practices with different patient populations (STAR-PU and ASTRO-PU).

| Column | Description |
|---|---|
| `month` | The month the statistics were recorded |
| `male_0_4` to `female_75_plus` | Count of patients by gender and age band (e.g. `male_0_4` is the number of male patients aged 0 to 4) |
| `total_list_size` | Total number of patients registered at the practice |
| `astro_pu_cost` | ASTRO-PU (Age, Sex, and Temporary Resident Originated Prescribing Units) cost weighting |
| `astro_pu_items` | ASTRO-PU items weighting |
| `star_pu` | STAR-PU (Specific Therapeutic group Age-sex Related Prescribing Units) weightings, stored as JSON |
| `pct_id` | Primary Care Trust identifier (e.g. 08G) |
| `practice` | Practice code (e.g. E86005) |

```{r}
practice_statistics_sample
```

## `regional_teams`

Names and open/close dates for regional teams.

| Column | Description |
|---|---|
| `code` | Regional team code (e.g. Y57) |
| `name` | Regional team name (e.g. SOUTH OF ENGLAND COMMISSIONING REGION) |
| `open_date` | Date the regional team opened |
| `close_date` | Date the regional team closed, `NA` if still open |

```{r}
regional_teams_sample
```

## How the tables join together

- `normalised_prescribing$practice` matches `practices$code`
- `normalised_prescribing$regional_team` matches `regional_teams$code`
- `practice_statistics$practice` matches `practices$code`

See the "Joining tables" section of `vignette("opr")` for a worked example.
