---
title: "Getting started with opr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with opr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The `get_*()` functions in `opr` return a query, not data.
The query runs in BigQuery, and nothing is downloaded until you call `collect()`.
This means you can keep adding `dplyr` steps like `filter()` and `select()` first, and only the (much smaller) result is downloaded.

```{r setup, message = FALSE}
library(opr)
library(dplyr)
```

The examples that query BigQuery are not run when this vignette is built.
To run them yourself you need access to the database, see `vignette("connect-bigquery")`.

## Connect

```{r, eval = FALSE}
con <- connect_bq()
```

## Get practices

`get_practices()` queries the `practices` table, with options to add descriptive labels for the `setting` and `status_code` columns and to filter by them.
For a description of the table see `vignette("database-schema")`, or look at the bundled five-row sample:

```{r}
practices_sample |>
  select(code, name, postcode, setting, status_code)
```

The labels come from lookup tables shipped with the package:

```{r}
practice_settings

practice_statuses
```

For example, to get all standard GP practices (setting code 4) with labels for setting and status codes:

```{r, eval = FALSE}
get_practices(
  con,
  add_setting_labels = TRUE,
  filter_setting = 4,
  add_status_code_labels = TRUE
) |>
  select(code, name, setting_label, status_code_label) |>
  collect()
```

To see the SQL that will run in BigQuery, use `show_query()`:

```{r, eval = FALSE}
get_practices(con, add_setting_labels = TRUE, filter_setting = 4) |>
  select(code, name, setting_label) |>
  show_query()
```

## Get normalised prescribing

`get_normalised_prescribing()` queries the `normalised_prescribing` table with optional filtering by BNF codes and prescribing month.
The full table is very large, so filter before collecting.
A sample of the table:

```{r}
normalised_prescribing_sample |>
  select(month, practice, bnf_code, bnf_name, items, quantity)
```

In this example we get prescribing data for two BNF codes between January and March 2023:

```{r, eval = FALSE}
get_normalised_prescribing(
  con,
  bnf_codes = c("0408010W0BBAKA1", "0601022B0AAASAS"),
  start_date = "2023-01-01",
  end_date = "2023-03-01"
) |>
  collect()
```

## Joining tables

Joins also run in BigQuery before anything is downloaded.
Here we join prescribing data to GP practices and filter for metformin:

```{r, eval = FALSE}
# Metformin 500mg modified-release tablets in March 2025
prescribing_query <- get_normalised_prescribing(
  con,
  bnf_codes = "0601022B0AAASAS",
  start_date = "2025-03-01",
  end_date = "2025-03-01"
) |>
  select(month, practice, bnf_code, bnf_name, items, quantity)

practices_query <- get_practices(con, filter_setting = 4) |>
  select(code, name, status_code)

prescribing_query |>
  inner_join(practices_query, by = c("practice" = "code")) |>
  collect()
```

## Other tables

The `get_*()` functions only cover the most commonly used tables.
The connection from `connect_bq()` works with any table in the dataset:

```{r, eval = FALSE}
dplyr::tbl(con, "practice_statistics")
```

And if you prefer writing SQL directly, that works too:

```{r, eval = FALSE}
DBI::dbGetQuery(con, "SELECT code, name FROM practices LIMIT 5")
```
