# Connect to BigQuery

## Get access

To get access to the OpenPrescribing BigQuery database write a message in the `#openprescribing` Slack channel and tag `bennett-admins`.
One of the admins will send you a JSON file with your BigQuery credentials in a private message. 
The file should look something like this:

```json
{
  "type": "ACCOUNT_TYPE",
  "project_id": "PROJECT_ID",
  "private_key_id": "PRIVATE_KEY_ID",
  "private_key": "PRIVATE_KEY",
  "client_email": "CLIENT_EMAIL",
  "client_id": "CLIENT_ID",
  "auth_uri": "AUTH_URI",
  "token_uri": "TOKEN_URI",
  "auth_provider_x509_cert_url": "AUTH_CERTIFICATE",
  "client_x509_cert_url": "CLIENT_CERTIFICATE",
  "universe_domain": "DOMAIN"
}
```
::: {.callout-important}
Never commit this file to GitHub or share it with anyone.
:::

## Store your credentials

You can store this file anywhere on your computer, but a common location is in a hidden folder in your home directory, for example:

**On macOS/Linux:**

```bash
~/.config/openprescribing/bq-credentials.json
```

**On Windows:**

```bash
C:/Users/YOUR_USERNAME/.config/openprescribing/bq-credentials.json
```

## Set up env variable {#sec-env-var}

Rather than hardcoding the path in your R scripts you can set an environment variable named `OP_CREDENTIALS` or similar.
To help you edit your `.Renviron` file and set the variable for all R sessions you can use the `usethis` package:

```{r}
#| label: edit-r-environ
#| eval: false

# Open your .Renviron file
usethis::edit_r_environ()
```

This will open your `.Renviron` file in your editor where you can add the location where you stored your credentials file, for example:

```bash
# macOS/Linux
OP_CREDENTIALS=~/.config/openprescribing/bq-credentials.json

# Windows
OP_CREDENTIALS=C:/Users/YOUR_USERNAME/.config/openprescribing/bq-credentials.json
```

**Note: You will probably need to restart your R session for R to know about your new environment variable.**

## Connet to BigQuery

The following code connects to the `hscic` dataset in the `ebmdatalab` project.
You will need to authenticate using the path to your credentials file stored in the `OP_CREDENTIALS` environment variable and might be asked to login with your Google account the first time you run this.

```{r}
#| label: connect-bq
#| eval: false

library(DBI)
library(bigrquery)

bq_auth(path = Sys.getenv("OP_CREDENTIALS"))

con <- dbConnect(
  bigrquery::bigquery(),
  project = "ebmdatalab",
  dataset = "hscic",
  billing = "ebmdatalab"
)
```
