# Database schema

```{r}
#| label: setup
#| include: false

library(DBI)
library(bigrquery)
library(dplyr)
library(here)
library(gt)
library(dm)

source(here("R", "connect-bq.R"))
source(here("R", "create-bq-desc-tbl.R"))
source(here("R", "tbl-dict.R"))

con <- connect_bq()
dbListTables(con)
```

## Main `hscic` tables

The `hscic` dataset has 194 different tables.
Here we describe the main tables used for research projects.
To see all the tables available in the `hscic` dataset, you can use:

```{r}
#| label: db-list-tables
#| eval: false

dbListTables(con)
```

### `normalised_prescribing` {#sec-desc-normalised-prescribing}

Normalised view of the main prescribing dataset, updated monthly.
This is the table you'll typically want to query rather than the raw prescribing table.

```{r}
#| label: desc-normalised-prescribing
#| echo: false
#| cache: true

normalised_prescribing <- tbl(con, "normalised_prescribing")

df_normalised_prescribing <- normalised_prescribing |>
  head(5) |>
  collect()

create_data_dict(
  df_normalised_prescribing,
  "normalised_prescribing"
)
```

### `regional_teams` {#sec-desc-regional-teams}

Open and close dates for regional teams.

```{r}
#| label: desc-regional-teams
#| echo: false
#| cache: true

regional_teams <- tbl(con, "regional_teams")

df_regional_teams <- regional_teams |>
  head(5) |>
  collect()

create_data_dict(
  df_regional_teams,
  "regional_teams"
)
```

### `practices` {#sec-desc-practices}

Includes all the practices in England, updated monthly.
Contains latest month's data only (but includes closed practices).
Practices with a `setting` of `4` are standard GP practices.

```{r}
#| label: desc-practices
#| echo: false
#| cache: true

practices <- tbl(con, "practices")

df_practices <- practices |>
  filter(setting == 4) |> 
  head(5) |>
  collect()

create_data_dict(
  df_practices,
  "practices"
)
```

### `practice_statistics` {#sec-desc-practice-statistics}

Total list size and list sizes stratified by gender and age group for each practice, updated monthly.
Contains latest month's data only. Also contains calculated population weightings related to prescribing (STAR-PU and ASTRO-PU).

```{r}
#| label: desc-practice-statistics
#| echo: false

practice_statistics <- tbl(con, "practice_statistics")

df_practice_statistics <- practice_statistics |>
  head(5) |>
  collect()

create_data_dict(
  df_practice_statistics,
  "practice_statistics"
)
```

## Relationship diagram?

I need someone to check this.
Could add some tables from the dmd dataset here (and above)?
Add description here.

```{r}
#| label: fig-hscic-relationships
#| echo: false
#| message: false
#| warning: false
#| cache: true
#| fig-height: 10
#| fig-width: 8
#| fig-cap: "Relationship diagram for `hscic` tables"

hscic_dm <- dm_from_con(
  con,
  schema = "hscic",
  table_names = c(
    "normalised_prescribing",
    "practices",
    "practice_statistics",
    "regional_teams"
  )
)

hscic_dm <- hscic_dm |>
  dm_add_pk(practices, code) |>
  dm_add_pk(regional_teams, code)

hscic_dm <- hscic_dm |>
  dm_add_fk(normalised_prescribing, practice, practices) |>
  dm_add_fk(normalised_prescribing, regional_team, regional_teams) |>
  dm_add_fk(practice_statistics, practice, practices)

hscic_dm |>
  dm_draw(view_type = "all")
```
