---
code-links:
  - text: connect-bq.R
    icon: file-code
    href: ../R/connect-bq.R
---

# Reusable functions

Here I'll demonstrate how `dbplyr` can be used to define reusable functions for:

1. **Connect to BigQuery**
2. **Query tables and perform common data cleaning tasks**
3. **Joining tables**

All functions here are based on the `DBI`, `bigrquery`, `dplyr` and `dbplyr` packages to connect to BigQuery and query the OpenPrescribing data in R.

```{r}
#| label: load-libraries
#| message: false
#| warning: false
library(DBI)
library(bigrquery)
library(dplyr)
library(dbplyr)
```

## Connect to BigQuery

We could write a simple function to connect to BigQuery that we can reuse in our R scripts.
This function assumes that you have set the `OP_CREDENTIALS` environment variable in @sec-env-var and we can define different datasets to connect to by changing the `dataset` argument (default is `hscic`).

```{r}
#| label: simple-connect-bq-function
connect_openprescribing <- function(dataset = "hscic") {
  # Get BigQuery credentials path from environment variable
  credentials_path <- Sys.getenv("OP_CREDENTIALS")

  # Authorise BigQuery
  bigrquery::bq_auth(path = credentials_path)

  # Connect to BigQuery
  DBI::dbConnect(
    bigrquery::bigquery(),
    project = "ebmdatalab",
    dataset = dataset,
    billing = "ebmdatalab"
  )
}
```

A slighly more robust and flexibly function could also have an option to work with different environment variables and return helpful error messages, see the example below.

```{r}
#| label: robust-connect-bq-function
#| file: ../R/connect-bq.R
#| code-fold: true
#| code-summary: Show more flexible and robust function to connect to BigQuery
```

## Data access functions

These functions provide a modular way to access and analyse OpenPrescribing data using `dbplyr`.
Each function returns a lazy query that can be composed with other `dplyr` functions.
For a more detailed overview see [Introduction to dbplyr](https://dbplyr.tidyverse.org/articles/dbplyr.html).

I'll demonstrate this how this could be useful for two commonly used tables: `practices` and `normalised_prescribing`.

### Get practices

To start we create a connection using the function we defined above:

```{r}
#| label: use-connect-bq-function
con <- connect_openprescribing()
```

For a description of the `practices` table see @sec-desc-practices.
Common data cleaning tasks when working with the `practices` table include adding filtering for specific pracitice settings and adding descriptive labels for setting and status codes.
We can write a function using `dplyr` syntax to make these tasks easier and reusable across projects.

```{r}
#| label: get-practices-function
#| code-fold: true
#| code-summary: Show `get_practices()` function to help access and tidy the `practices` table

get_practices <- function(
  con,
  add_setting_labels = FALSE,
  filter_setting = NULL,
  add_status_code_labels = FALSE,
  filter_status_code = NULL
) {

  practices_query <- dplyr::tbl(con, "practices")

  # Add status code description labels
  if (add_status_code_labels) {
    practices_query <- practices_query |>
      dplyr::mutate(
        status_code_label = dplyr::case_when(
          status_code == "A" ~ "Active",
          status_code == "B" ~ "Retired",
          status_code == "C" ~ "Closed",
          status_code == "D" ~ "Dormant",
          status_code == "P" ~ "Proposed",
          status_code == "U" ~ "Unknown",
          .default = "Unknown"
        )
      )
  }

  # Add setting description labels
  if (add_setting_labels) {
    practices_query <- practices_query |>
      dplyr::mutate(
        setting_label = dplyr::case_when(
          setting == 0  ~ "Other",
          setting == 1  ~ "WIC Practice",
          setting == 2  ~ "OOH Practice",
          setting == 3  ~ "WIC + OOH Practice",
          setting == 4  ~ "GP Practice",
          setting == 8  ~ "Public Health Service",
          setting == 9  ~ "Community Health Service",
          setting == 10 ~ "Hospital Service",
          setting == 11 ~ "Optometry Service",
          setting == 12 ~ "Urgent & Emergency Care",
          setting == 13 ~ "Hospice",
          setting == 14 ~ "Care Home / Nursing Home",
          setting == 15 ~ "Border Force",
          setting == 16 ~ "Young Offender Institution",
          setting == 17 ~ "Secure Training Centre",
          setting == 18 ~ "Secure Childrens Home",
          setting == 19 ~ "Immigration Removal Centre",
          setting == 20 ~ "Court",
          setting == 21 ~ "Police Custody",
          setting == 22 ~ "Sexual Assault Referral Centre (SARC)",
          setting == 24 ~ "Other – Justice Estate",
          setting == 25 ~ "Prison",
          setting == 26 ~ "PCN",
          setting == 27 ~ "Independent Pharmacy",
          .default = "Unknown"
        )
      )
  }

  # Filter by setting codes
  if (!is.null(filter_setting)) {
    practices_query <- practices_query |>
      dplyr::filter(setting %in% filter_setting)
  }

  # Filter by setting codes
  if (!is.null(filter_status_code)) {
    practices_query <- practices_query |>
      dplyr::filter(status_code %in% filter_status_code)
  }

  practices_query
}
```

Now we can use the connection `con` and the function `get_practices()` to get practices data.
Here are the first 5 rows of some columns we can select using the `select()` function from `dplyr`, unfiltered and without labels:

```{r}
#| label: get-practices-function-example-1
get_practices(con) |>
  select(code, name, postcode, setting, status_code) |>
  head(5)
```


We can specify to add labels for the setting codes and filter to only include standard GP practices (setting code 4). 

```{r}
#| label: get-practices-function-example-2
get_practices(con, add_setting_labels = TRUE, filter_setting = 4) |> 
  select(code, name, postcode, setting, setting_label, status_code) |> 
  head(5)
```

Also add labels for the status codes:

```{r}
#| label: get-practices-function-example-3
get_practices(con, add_setting_labels = TRUE, filter_setting = 4,
              add_status_code_labels = TRUE) |> 
  select(code, name, postcode, setting, setting_label, status_code, status_code_label) |> 
  head(5)
```

To inspect the SQL query that `dbplyr` generates for this last example we can use the `show_query()` function:

```{r}
#| label: get-practices-function-show-sql
get_practices(con, add_setting_labels = TRUE, filter_setting = 4,
              add_status_code_labels = TRUE) |> 
  select(code, name, postcode, setting, setting_label, status_code, status_code_label) |> 
  head(5) |> 
  show_query()
```

### Get normalised prescribing

For a description of the `normalised_prescribing` table see @sec-desc-normalised-prescribing.
This function returns normalised prescribing data with optional filtering by BNF codes and date range.

```{r}
#| label: get-normalised-prescribing-function
#| code-fold: true
#| code-summary: Show `get_normalised_prescribing()` function to help access and tidy the `normalised_prescribing` table

get_normalised_prescribing <- function(
  con,
  bnf_codes = NULL,
  start_date = NULL,
  end_date = NULL) {
  
  np_query <- dplyr::tbl(con, "normalised_prescribing")

  # Filter by BNF codes if provided
  if (!is.null(bnf_codes)) {
    np_query <- np_query |>
      dplyr::filter(bnf_code %in% !!bnf_codes)
  }

  # Filter by start date if provided
  if (!is.null(start_date)) {
    np_query <- np_query |>
      dplyr::filter(month >= !!start_date)
  }

  # Filter by end date if provided
  if (!is.null(end_date)) {
    np_query <- np_query |>
      dplyr::filter(month <= !!end_date)
  }

  np_query
}
```

In this example we get normalised prescribing data for two BNF codes between January and March 2023.

```{r}
#| label: get-normalised-prescribing-example-1
get_normalised_prescribing(
  con, 
  bnf_codes = c("0408010W0BBAKA1", "0601022B0AAASAS"),
  start_date = "2023-01-01",
  end_date = "2023-03-01"
  ) |> 
  filter(practice == "D83033") |> 
  head(5)
```

## Joining tables

```{r}
#| label: join-practices-normalised-prescribing
 
# Create query for normalised prescribing 
normalised_prescribing_query <- get_normalised_prescribing(con) |> 
  select(month, practice, bnf_code, bnf_name, items, quantity)

# Create query for practices 
practices_query <- get_practices(con, add_setting_labels = TRUE, filter_setting = 4) |> 
  select(code, name, setting, setting_label, status_code)

# Join by practice / code
joined_query <- normalised_prescribing_query |> 
  inner_join(practices_query, by = c("practice" = "code"))
```

We can now filter this query futher using dplyr syntax, e.g., from a predefined list of BNF codes:

```{r}
#| label: filter-joined-query

# Define a list of BNF codes
# Metformin 500mg modified-release tablets
metformin_bnf_codes <- c("0601022B0AAASAS")

joined_query |> 
  filter(bnf_code %in% metformin_bnf_codes) |> 
  filter(month == "2025-03-01")
```

## Writing SQL with dbplyr

See [Writing SQL with dbplyr](https://dbplyr.tidyverse.org/articles/sql.html).
